# R Code for edX Competition on *"Analytics Edge"* DataScience Course
==================================================================

#### Author: Fahad Usman
#### Date: 21 April 2014

## The Problem Statement: Learn what predicts happiness by using informal polling questions. 

What predicts happiness? In this competition, you'll be using data from Show of Hands, an informal polling platform for use on mobile devices and the web, to see what aspects and characteristics of people's lives predict happiness.

Show of Hands has been downloaded over 300,000 times across Apple and Android app stores, and users have cast more than 75 million votes. In this problem, we'll use data from thousands of users and one hundred different questions to see which responses predict happiness.

## Acknowledgements

This competition is brought to you by 15.071x, edX, and Show of Hands.

## File descriptions

Here is a description of the files you have been provided for this competition:

- **train.csv** - the training set of data that you should use to build your models
- **test.csv**  - the test set that you will be evaluated on. It contains all of the independent variables, but not the dependent variable.
- **sampleSubmission.csv** - a sample submission file in the correct format.
- **Questions.pdf** - the question test corresponding to each of the question codes, as well as the possible answers.

## Data fields

- **UserID** - an anonymous id unique to a given user
- **YOB** - the year of birth of the user
- **Gender** - the gender of the user, either Male, Female, or not provided
- **Income** - the household income of the user. Either not provided, or one of "under $25,000", "$25,001 - $50,000", "$50,000 - $74,999", "$75,000 - $100,000", "$100,001 - $150,000", or "over $150,000".
- **HouseholdStatus** - the household status of the user. Either not provided, or one of "Domestic Partners (no kids)", "Domestic Partners (w/kids)", "Married (no kids)", "Married (w/kids)", "Single (no kids)", or "Single (w/kids)".
- **EducationLevel** - the education level of the user. Either not provided, or one of "Current K-12", "High School Diploma", "Current Undergraduate", "Associate's Degree", "Bachelor's Degree", "Master's Degree", or "Doctoral Degree".
- **Party** - the political party of the user. Either not provided, or one of "Democrat", "Republican", "Independent", "Libertarian", or "Other".
- **Happy** - a binary variable, with value 1 if the user said they were happy, and with value 0 if the user said that were neutral or not happy. This is the variable you are trying to predict.
- Q124742, Q124122, . . . , Q96024 - 101 different questions that the users were asked on Show of Hands. If the user didn't answer the question, there is a blank. For information about the question text and possible answers, see the file Questions.pdf.
- **votes** - the total number of questions that the user responded to, out of the 101 questions included in the data set (this count does not include the happiness question).


Let's begin by reading in the train file:

```{r, include=FALSE}
library(markdown)
library(knitr)
```

```{r}
happinessTrain = read.csv("C:/Users/Fahad/Documents/R Projects/edXCompetition/Data/train.csv")
```

Always check the data structure first:

```{r}
str(happinessTrain)
summary(happinessTrain)
```

So we have 4619 observations and 110 columns/variables,